---
title: "FinalProjectR"
author: "Teri Massey /Karen Weigandt"
date: "July 26, 2015"
output: html_document
---


Install necessary packages:
```{r}
install.packages("devtools", dependencies = TRUE, repos = "http://lib.stat.cmu.edu/R/CRAN/")
install.packages("ggplot2", dependencies = TRUE, repos = "http://lib.stat.cmu.edu/R/CRAN/")

```

Get RCurl on board so csv files can be retrieved, get gglot2 on board

```{r}
library(RCurl)
library(ggplot2)

```


Retrieve data file from github repository
```{r}
ten_eleven_comp <- getURL("https://raw.githubusercontent.com/karenweigandt/cuny-summer-bridge/master/Libraries_Annual_Statistics_Comparison_2010-2011.csv")
eleven_coll_stats <- getURL("https://raw.githubusercontent.com/karenweigandt/cuny-summer-bridge/master/Libraries_Collection_Statistics_2011_2.csv")
five_twelve_mat_inv <- getURL("https://raw.githubusercontent.com/karenweigandt/cuny-summer-bridge/master/Libraries_Material_Inventory_2005_-_2012.csv")

```

Read data file
```{r}
ten_eleven_comp_csv <- read.csv(text = ten_eleven_comp, header = TRUE, stringsAsFactors = FALSE)
eleven_coll_stats_csv <- read.csv(text = eleven_coll_stats, header = TRUE, stringsAsFactors = FALSE)
five_twelve_mat_inv_csv <- read.csv(text = five_twelve_mat_inv, header = TRUE, stringsAsFactors = FALSE)

```


Look at first six rows of data
```{r}
head(ten_eleven_comp_csv)
head(eleven_coll_stats_csv)
head(five_twelve_mat_inv_csv)

```

Subset material inventory data to only include 2011
```{r}
eleven_mat_inv <- subset(five_twelve_mat_inv_csv, Year == 2011)
#head(eleven_mat_inv)

```

Subset 2011 material inventory to books on the island of Oahu
```{r}
oahu_eleven_book_inv <- subset(eleven_mat_inv, ISLAND == 'Oahu' & INVENTORY.DESCRIPTOR == 'Library Materials Books')

#oahu_eleven_book_inv
```


Histogram for Books in Hawaii libraries
```{r}
options(sscipen = 3)
qplot(eleven_coll_stats_csv$BOOK, 
      geom = "histogram", 
      binwidth = 25000,
      col = I("red"),
      fill = I("blue"),
      main = "Books in Hawaii Libraries",
      xlab = "Number of Books",
      ylab = "Number of Libraries"
      )

```

Basic Statistics for number of books in Hawaii libraries
```{r}
sprintf("Average number of books in Hawaii libraries is %s", round(mean(eleven_coll_stats_csv$BOOK)))
sprintf("Average standard deviation for the number of books in Hawaii libraries is %s", round(sd(eleven_coll_stats_csv$BOOK)))
sprintf("Median number of books in Hawaii libraries is %s", round(median(eleven_coll_stats_csv$BOOK)))

```


Quantile for the number of books in Hawaii libraries
```{r}
qb <- quantile(eleven_coll_stats_csv$BOOK)
qb

```


Boxplot for number of books in Hawaii libraries
```{r}
options(scipen = 3)
boxplot(eleven_coll_stats_csv$BOOK, main = "Boxplot: Books in Hawaii Libraries")

```


```{r}
eleven_patron_circ <- ten_eleven_comp_csv[c(-2, -4, -6, -7, -9, -10, -12,-13, -14, -15)]

#head(eleven_patron_circ)

eleven_oahu_patron_circ <- subset(eleven_patron_circ, Island == 'Oahu')
 
#eleven_oahu_patron_circ

eleven_kauai_patron_circ <- subset(eleven_patron_circ, Island == 'Kauai')

#head(eleven_kauai_patron_circ)
```

Plot holdings vs. books

```{r}
qplot(TOTAL, BOOK,
      data = eleven_coll_stats_csv, 
      geom = "point", 
      main = "Year 2011: Holdings vs. Number of Books for All Hawaii Libraries",
      xlab = "Number of Holdings",
      ylab = "Number of Books"
      )

```




Display a scatter plot of Holdings vs. Circulation for Hawaii libraries or 2011

```{r}
qplot(Holdings.FY.2011, Circulation.Activity.FY.2011,
      data = eleven_patron_circ, 
      geom = "point", 
      main = "Year 2011: Holdings vs. Circulation in Hawaii Libraries",
      xlab = "Number of Holdings",
      ylab = "Circulation Volume"
      )

```



Display a scatter plot of Holdings vs. Circulation for Oahu libraries or 2011

```{r}
qplot(Holdings.FY.2011, Circulation.Activity.FY.2011,
      data = eleven_oahu_patron_circ, 
      geom = "point", 
      main = "Year 2011: Holdings vs. Circulation in Oahu Libraries",
      xlab = "Number of Holdings",
      ylab = "Circulation Volume"
      )


```


Display a scatterplot of Borrowers vs. Circulation for Oahu libraries or 2011

```{r}
options(scipen = 3)
qplot(Registered.Borrowers.FY.2011, Circulation.Activity.FY.2011,
      data = eleven_oahu_patron_circ, 
      geom = "point", 
      main = "Year 2011: Borrowers vs. Circulation in Oahu Libraries",
      xlab = "Number of Borrowers",
      ylab = "Circulation Volume"
      )

```

Display a scatterplot of Borrowers vs. Holdings for Oahu libraries or 2011

```{r}
options(scipen = 3)
qplot(Registered.Borrowers.FY.2011, Holdings.FY.2011,
      data = eleven_oahu_patron_circ, 
      geom = "point", 
      main = "Year 2011: Borrowers vs. Holdings in Oahu Libraries",
      xlab = "Number of Borrowers",
      ylab = "Number of Holdings"
      )

```



Display a scatter plot of Holdings vs. Circulation for Kauai libraries or 2011

```{r}
qplot(Holdings.FY.2011, Circulation.Activity.FY.2011,
      data = eleven_kauai_patron_circ, 
      geom = "point", 
      main = "Year 2011: Holdings vs. Circulation in Kauai Libraries",
      xlab = "Number of Holdings",
      ylab = "Circulation Volume"
      )


```


Display a scatterplot of Borrowers vs. Circulation for Kauai libraries or 2011

```{r}
options(scipen = 3)
qplot(Registered.Borrowers.FY.2011, Circulation.Activity.FY.2011,
      data = eleven_kauai_patron_circ, 
      geom = "point", 
      main = "Year 2011: Borrowers vs. Circulation in Kauai Libraries",
      xlab = "Number of Borrowers",
      ylab = "Circulation Volume"
      )

```


Display a scatterplot of Borrowers vs. Holdings for Kauai libraries or 2011

```{r}
options(scipen = 3)
qplot(Registered.Borrowers.FY.2011, Holdings.FY.2011,
      data = eleven_kauai_patron_circ, 
      geom = "point", 
      main = "Year 2011: Borrowers vs. Holdings in Kauai Libraries",
      xlab = "Number of Borrowers",
      ylab = "Number of Holdings"
      )

```


COnclusion:

The information and visualizations presented herein allow us to easily ascertain some information about the libraries inn Hawaii.  First, the histogram shows that the majority of Hawaiian libraries contain less than 100,000 books

