---
title: "Assignment4RegularExpressionsMassey"
output: html_document
---

```{r setup, include=TRUE}
library(stringr)
library(stringr)
library(XML)
library(maps)
```

3. Copy the introductory example. The vector name stores the extracted names.

```{r}
#raw.data = input data
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"

#Search for a capital letter followed by lowercase letters with two more minimal characters no min.
#extract names
name <- unlist(str_extract_all(raw.data, "[[A-Z][a-z]+., ]{2,}"))
name

```

(a) Use the tools of this chapter to rearrange the vector so that all elements conform to
the standard first_name last_name.
```{r}

sp_name <-str_split(name,",")
sp_name
```
```{r}
for (i in 1:length(sp_name)){
  if (length(sp_name[[i]]) > 1) {
    temp <- sp_name[[i]][1]
    sp_name[[i]][1] <- sp_name[[i]][2]
    sp_name[[i]][2] <- temp
  }
}
sp_name
```

(b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and
Dr.).
```{r}
title_name <- str_detect(name,"Rev.|Dr.|Mr.|Ms.|Mrs.")
booltn <- data.frame (name, title_name)
booltn



```

(c) Construct a logical vector indicating whether a character has a second name.
```{r}
middle_name <- str_detect(name," [A-Z]\\.")
boolmn <- data.frame (name, middle_name)
boolmn
```

7. Consider the string <title>+++BREAKING NEWS+++</title>. We would like to
extract the first HTML tag. To do so we write the regular expression <.+>. Explain why
this fails and correct the expression.
```{r}
tag_str <- "<title>+++BREAKING NEWS+++</title>"


wrong_tag <-str_extract(tag_str, "<.+>")
wrong_tag
```

```{r}
#the regex
right_tag <-str_extract(tag_str, "<[[:alpha:]]+>")
right_tag
```


8. Consider the string (5-3)2=52-253+3^2 conforms to the binomial theorem. We would like to extract the formula in the string. To do so we write the regular expression [^0-9=+*()]+. Explain why this fails and correct the expression.


```{r}
binofo <- "(5-3)^2=5^2-2*5*3+3^2"


fix_form <-str_extract(binofo, "[^0-9=+*()]+")
fix_form

#the regex
new_form <-str_extract(binofo, "[\\^0-9-=+*()]+")
new_form
```


9. The following code hides a secret message. Crack it with R and regular expressions.
Hint: Some of the characters are more revealing than others! The code snippet is also
available in the materials at www.r-datacollection.com.
clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr


```{r}
tempstr <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

secret <-unlist(str_extract_all(tempstr, "[[:upper:].!]"))

secret
```
